Deep triplet network adopting the kernel and the range space learning for Wi-Fi signature verification

### 2019-07-08 190410_Paper_ACPR/paper_tcnn_kar.tex
paper_acpr: methods

\subsection{ConvNet with triplet loss}

% description of siamese net
In \cite{bromley1994signature}, LeCun et al. Introduced Siamese nets as parts of their handwritten signature verification system. 
A siamese neural network consists of twin networks which accept distinct inputs but are joined by an energy function at the top.\cite{koch2015siamese}
% differences in feature extractor
They proposed a feature extractor based on Time Delay Neural Networks\cite{lang1990time} and feature matcher based on cosine distance of feature vector. Apart from this work based on ConvNet as feature extractor and L1 distance as feature matcher.
In this paper, we propose a for signature verification system that is applicable for WIFI CSI signal, which is more complex and larger than handwritten images. 
% advantages of using Convnet filter
Based on Convnets as feature extractor, we are able to make feature vector from CSI signal reflecting local connectivity between signals at a near frequency range and closer measurement time. 
The proposed method achieved better performance because it is applicable to all points of CSI signals. This is due to the weight sharing the property of the Convnet filter.

\section{Proposed System}

In this section, we propose a direction-free identify verification system based on the Wi-Fi based in-air handwritten signature (will be called Wi-Fi signature signals hereafter). An overview of the proposed system utilizing the ConvNet with triplet loss is shown in Fig.1.
Essentially, the Wi-Fi signature signals are preprocessed to create the input data for our network. Subsequently, the ConvNet structure is trained to minimize distance from anchor to positive input while maximize distance from anchor to negative inputs. The following subsections detail the data preprocessing and the proposed method.

@@

\subsection{Data Preprocessing}

Since every Wi-Fi signature signal has different data size, we firstly adopted the gradient operation with respect to the time instance to measure the short time energy. Data points with the highest short-time energy within the time period are then manually selected as the starting and the ending points of the in-air signature action. Subsequently, the Fast Fourier Transform based re-sampling method \cite{moon2017air} is implemented to unify the length of the signals. As a result, three-dimensional Wi-Fi signature signals with unified data size are obtained as the input of the ConvNet structure in the network.

% Methods 
% Section: Feature Extraction
\subsection{ConvNets with triplet loss}

@@

In our proposed networks, we utilize the $L_2$ distance to calculate the triplet loss. For the $i_{th}$ signal as anchor input $\mathbf{I}_{i,anc}$, positive input $\mathbf{I}_{i,pos}$ is belong to the same class for anchor input while negative input $\mathbf{I}_{i,neg}$ is from another class.
Let $\mathbf{x}_{i,anc}\in{\mathrm{R}}^{d\times1}$, $\mathbf{x}_{i,pos}\in{\mathrm{R}}^{d\times1}$ and $\mathbf{x}_{i,neg}\in{\mathrm{R}}^{d\times1}$ be three feature vectors extracted from the ConvNet structure. Then the loss for the triplet inputs can be calculated as follows:
\begin{equation}
loss = \sum_i^N { \left[ {\left\| {{\mathbf{x}_{i,anc}} - {\mathbf{x}_{i,pos}}} \right\|_2^2} -
{\left\| {{\mathbf{x}_{i,anc}} - {\mathbf{x}_{i,neg}}} \right\|_2^2}  + \alpha \right]_+},
\end{equation} 
where $\alpha$ is the margin to make distance between positive and negative pairs, N is size of the mini-batch.

% KAR learning
\subsection{Triplet selection by the kernel and the range space learning}

% methods of KAR learning
since we don't know which is the hard sample before training entire network, we make multilayer feedforward neural network instead to select the hard positive and the hard negative.
we adopted gradient-free learning to train this network. by the kernel and the range (KAR) space projection learning \cite{toh2018learning,toh2018gradient}.

% training KARnet
first, let the training dataset $\mathbf{X}\in{\mathrm{R}}^{m \times (n+1)}$ and one-hot encoded target $\mathbf{Y}\in{\mathrm{R}}^{m \times n}$.

set multilayer neural network structure 
\begin{equation}
\mathbf{Y} = \sigma\left(\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(i-1)}\right)\right]\mathbf{W}_{i}\right),
\end{equation}

the weight matrix $\mathbf{W}_{1}\dots\mathbf{W}_{i}$ can be separated into weights and bias term,
$\mathbf{W}_{2}\dots\mathbf{W}_{i}$ = 
$\begin{bmatrix}
\mathbf{w}_{2}^{T}\\\mathnormal{W}_{2}
\end{bmatrix}
\dots
\begin{bmatrix}
\mathbf{w}_{i}^{T}\\\mathnormal{W}_{i}
\end{bmatrix}$

%\begin{equation}
\begin{multline*}
\Rightarrow
\biggl[\sigma^{-1}\biggl(
\dots
\biggl[\sigma^{-1}\biggl(
    \left[\sigma^{-1}\left(\mathbf{Y}\right)-\mathbf{1}\cdot\mathbf{w}_{i}^{T}\right]\mathnormal{W}_{i}^\dagger
\biggr)\\
- \mathbf{1}\cdot\mathbf{w}_{(i-1)}^{T}\biggr]\mathnormal{W}_{(i-1)}^{\dagger}
\dots\biggr) 
- \mathbf{1}\cdot\mathbf{w}_{2}^{T}\biggr]\mathnormal{W}_{2}^{\dagger} 
= \sigma\left(\mathbf{X}\mathbf{W}_{1}\right)
\end{multline*}
%\end{equation}

\begin{multline*}
\Rightarrow
\mathbf{X}^{\dagger}\sigma^{-1}\biggl(
\biggl[\sigma^{-1}\biggl(
\dots
\biggl[\sigma^{-1}\biggl(
    \left[\sigma^{-1}\left(\mathbf{Y}\right)-\mathbf{1}\cdot\mathbf{w}_{i}^{T}\right]\mathnormal{W}_{i}^\dagger
\biggr)\\
- \mathbf{1}\cdot\mathbf{w}_{(i-1)}^{T}\biggr]\mathnormal{W}_{(i-1)}^{\dagger}
\dots\biggr)
- \mathbf{1}\cdot\mathbf{w}_{2}^{T}\biggr]\mathnormal{W}_{2}^{\dagger}
= \mathbf{W}_{1}
\end{multline*}

\begin{multline*}
\Rightarrow
\left(\sigma\left(\mathbf{X}\mathbf{W}_{1}\right)\right)^{\dagger}
\biggl(\dots
\biggl[\sigma^{-1}\biggl(
    \left[\sigma^{-1}\left(\mathbf{Y}\right)-\mathbf{1}\cdot\mathbf{w}_{i}^{T}\right]\mathnormal{W}_{i}^\dagger
\biggr)
- \mathbf{1}\cdot\mathbf{w}_{(i-1)}^{T}\biggr]\mathnormal{W}_{(i-1)}^{\dagger}
\dots\biggr)\\ 
= \mathbf{W}_{2}
\end{multline*}

\section{Experiments}
% Dataset
\subsubsection{Dataset}
 To evaluate validation performance of proposed system, Wi-Fi CSI signature dataset from \cite{moon2017air} was used.
 % description
 we utilized 2000 Wi-Fi CSI signature signals (4 directions $\times$ 50 identities $\times$ 10 samples) which is dimension of (500 packets $\times$ 30 subcarriers $\times$ 6 antennas).

@@

% Network Learning
\subsubsection{Loss and Backpropagation}
 We impose a regularized cross-entropy objective on our binary classifier.
This objective is combined with standard backpropagation algorithm, where the gradient is additive across the twin networks due to the tied weights.
 We initialized all network weights in the convolutional layers from a normal distribution with zero-mean and a standard deviation of 10−2. Biases were also initialized from a normal distribution, but with mean 0.5 and standard deviation 10−2.

>>>>

\subsection{the kernel and the range space learning}

Multilayer feedforward neural networks has been widely used to feature extractor and classifier.
For traning the feedforward multilayer networks, non-gradient learning method by using linear matrix equation has invented recently. \cite{wang2018review}

This model avoids the problems of the deep learning model such as vanishing gradient and local minima. And except for the number of layers and neurons used in the network, there are few parameters for the network configuration.

More recently, a new learning framework has been developed to train the weights of the deep networks by a series of kernel and range space manipulations \cite{toh2018gradient,toh2018learning}.
This method allows to quickly learn the weights of the deep network with little  system resources.
we adopted this method to mining negative samples for the anker signal.

%% Methods
\section{Proposed System}

In this section, we propose a direction-free identify verification system based on the Wi-Fi based in-air handwritten signature (will be called Wi-Fi signature signals hereafter). An overview of the proposed system utilizing the ConvNet \cite{lecun1998gradient} and the kernel and the range space projection learning (KAR space learning) is shown in Fig.1.
Essentially, the Wi-Fi signature signals are preprocessed to create the input data for our network. Subsequently, the training dataset is moved to KAR space learning. it mines the hard positve and negative samples for given anchor signal from training dataset whtout using gradient.
ConvNet structure extract feature vectors from given triplet using convnet filters.
ConvNet training is achived by triplet loss, which is L2 distance of given feature vectors of the triplet. 
The following subsections detail the proposed method.

@@

% Methods: Triplet loss
\subsection{Triplet loss}

Our proposed networks utilizes the $L_2$ distance to calculate the triplet loss. For the $i_{th}$ signal as anchor input $\mathbf{I}_{i,anc}$, positive input $\mathbf{I}_{i,pos}$ is belong to the same class for anchor input while negative input $\mathbf{I}_{i,neg}$ is from another class.
Let $\mathbf{x}_{i,anc}\in{\mathrm{R}}^{d\times1}$, $\mathbf{x}_{i,pos}\in{\mathrm{R}}^{d\times1}$ and $\mathbf{x}_{i,neg}\in{\mathrm{R}}^{d\times1}$ be three feature vectors extracted from the ConvNet structure. Then the loss for the triplet inputs can be calculated as follows:
\begin{equation}
loss = \sum_i^N { \left[ {\left\| {{\mathbf{x}_{i,anc}} - {\mathbf{x}_{i,pos}}} \right\|_2^2} -
{\left\| {{\mathbf{x}_{i,anc}} - {\mathbf{x}_{i,neg}}} \right\|_2^2}  + \alpha \right]_+},
\end{equation} 
where $\alpha$ is the margin to make distance between positive and negative pairs, N is size of the mini-batch.
+ sign means greater between zero or calculated loss.

% Methods: KAR learning

\subsection{Triplet mining by the kernel and the range space learning}

% methods of KAR learning
since we don't know which is the hard sample before training entire network, we make multilayer feedforward neural network to select the hard positive and the hard negative.
we adopted gradient-free, the kernel and the range (KAR) space projection learning to train multilayer feedforward neural network. \cite{toh2018learning,toh2018gradient}.

% training KARnet
first, multilayer neural network structure is shown below:

\begin{equation}
\mathbf{G} = \sigma\left(\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(i-1)}\right)\right]\mathbf{W}_{i}\right),
\end{equation}

let the training dataset $\mathbf{X}\in{\mathrm{R}}^{m \times (n+1)}$ and $\mathbf{G}\in{\mathrm{R}}^{m \times n}$ is network outputs.

network learning is archived using one-hot encoded target $\mathbf{Y}\in{\mathrm{R}}^{m \times n}$ instead of $\mathbf{G}$.

the weight matrix $\mathbf{W}_{1}\dots\mathbf{W}_{i}$ can be separated into weights and bias term,
$\mathbf{W}_{2}\dots\mathbf{W}_{i}$ = 
$\begin{bmatrix}
\mathbf{w}_{2}^{T}\\\mathnormal{W}_{2}
\end{bmatrix}
\dots
\begin{bmatrix}
\mathbf{w}_{i}^{T}\\\mathnormal{W}_{i}
\end{bmatrix}$.

%\begin{multline*}
\begin{equation}
\begin{aligned}
\Rightarrow
\biggl[\sigma^{-1}\biggl(
\dots
\biggl[\sigma^{-1}\biggl(
    \left[\sigma^{-1}\left(\mathbf{Y}\right)-\mathbf{1}\cdot\mathbf{w}_{i}^{T}\right]\mathnormal{W}_{i}^\dagger
\biggr)
- \mathbf{1}\cdot\mathbf{w}_{(i-1)}^{T}\biggr]\mathnormal{W}_{(i-1)}^{\dagger}
\dots\biggr)\\
- \mathbf{1}\cdot\mathbf{w}_{2}^{T}\biggr]\mathnormal{W}_{2}^{\dagger} 
= \sigma\left(\mathbf{X}\mathbf{W}_{1}\right)
%\end{multline*}
\end{aligned}
\end{equation}

%\begin{multline*}
\begin{equation}
\begin{aligned}
\Rightarrow
\mathbf{X}^{\dagger}\sigma^{-1}\biggl(
\biggl[\sigma^{-1}\biggl(
\dots
\biggl[\sigma^{-1}\biggl(
    \left[\sigma^{-1}\left(\mathbf{Y}\right)-\mathbf{1}\cdot\mathbf{w}_{i}^{T}\right]\mathnormal{W}_{i}^\dagger
\biggr)
- \mathbf{1}\cdot\mathbf{w}_{(i-1)}^{T}\biggr]\mathnormal{W}_{(i-1)}^{\dagger}
\dots\biggr)\\
- \mathbf{1}\cdot\mathbf{w}_{2}^{T}\biggr]\mathnormal{W}_{2}^{\dagger}
= \mathbf{W}_{1}
\end{aligned}
\end{equation}
%\end{multline*}

\begin{equation}
\begin{aligned}
%\begin{multline*}
\Rightarrow
\left(\sigma\left(\mathbf{X}\mathbf{W}_{1}\right)\right)^{\dagger}
\biggl(\dots
\biggl[\sigma^{-1}\biggl(
    \left[\sigma^{-1}\left(\mathbf{Y}\right)-\mathbf{1}\cdot\mathbf{w}_{i}^{T}\right]\mathnormal{W}_{i}^\dagger
\biggr)
- \mathbf{1}\cdot\mathbf{w}_{(i-1)}^{T}\biggr]\mathnormal{W}_{(i-1)}^{\dagger}
\dots\biggr)\\
= \mathbf{W}_{2}
%\end{multline*}
\end{aligned}
\end{equation}

when anker signal is set, negative samples can be mined if the distance of network output $\mathbf{G}$ is above the threshold.

% Methods: ConvNets
\subsection{ConvNet Structures}

@@

\section{Experiments}

% Dataset
\subsubsection{Dataset}
 To evaluate validation performance of proposed system, Wi-Fi CSI signature dataset from \cite{moon2017air} was used.
 % pre-processing
 Since every Wi-Fi signature signal has different data size, we firstly adopted the gradient operation with respect to the time instance to measure the short time energy. Data points with the highest short-time energy within the time period are then manually selected as the starting and the ending points of the in-air signature action. Subsequently, the Fast Fourier Transform based re-sampling method \cite{moon2017air} is implemented to unify the length of the signals. As a result, three-dimensional Wi-Fi signature signals with unified data size are obtained as the input of the ConvNet structure in the network.
 % description
 we utilized 2000 Wi-Fi CSI signature signals (4 directions $\times$ 50 identities $\times$ 10 samples) which is dimension of (500 packets $\times$ 30 subcarriers $\times$ 6 antennas).

@@

% Network Learning
\subsubsection{ConvNet structure}
 We impose a triplet loss objective on our classifier.
This objective is combined with standard backpropagation algorithm, where the gradient is additive across the twin networks due to the tied weights.
 We initialized all network weights in the convolutional layers from a normal distribution with zero-mean and a standard deviation of 10−2. Biases were also initialized from a normal distribution, but with mean 0.5 and standard deviation 10−2.
